Choose a License

### 2024-09-12 choose_license.qmd
Minor stylistic improvements

::: {#cau-exclusive-rights .callout-caution}
### Publishing Without Exclusive Rights

When you have your article published under an open access license, some publishers still demand an exclusive publishing and distribution license that goes further than the open access license available to readers of the journal [@Rumsey2022]. However, technically this is not required for publishers to provide their services [@Suber2022] and authors should oppose this practice. If your chosen publisher insists on an exclusive license, you may at least retain the copyright for your figures -- follow the guide by @Elson2016 to learn how to do that.

@@

For example, the `apaquarto` extension that you included in your project contains works by others. You need to indicate its license so that others know what they are allowed to do -- and, of course, you need to comply with its terms yourself, such as retaining the copyright notice.^[In this particular case, `apaquarto` is licensed under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/), which does not require you to retain a copyright notice. In fact, it's an extremely permissive license.] In contrast, if it were not for this tutorial, the manuscript would contain your own work and you would need to indicate under which license you provide it to others.

@@

:::: {}
- [CC licenses](https://creativecommons.org/faq/#how-do-i-apply-a-creative-commons-license-to-my-material)
- [ODbL](https://opendatacommons.org/licenses/odbl/)
::::
:::

@@

In the simplest case, one just creates a file called `LICENSE.txt` where the full text of the license is copied verbatim. This is a practice propagated by GitHub, who provides instructions for many licenses via [ChooseALicense.com](https://choosealicense.com/licenses/). However, if the project is not completely covered by one single license, this practice may become unwieldy. For example, if a project contains different types of works by different authors, the `LICENSE.txt` needs to detail which file is covered by which license(s), along with any copyright notices.

@@

Every major free/open license has a unique [SPDX identifier](https://spdx.org/licenses/) which allows communicating the license choice unequivocally. We will be using that to indicate the license for every file in your project, along with the year of publication and the copyright holder. To do this, we add a comment to the beginning of every file and add the two tags `SPDX-FileCopyrightText` and `SPDX-License-Identifier`. How this works depends on the file type, as the syntax for a comment varies.

For example, if you previously created the file `create_data_dictionary.R`, you can now add the following comment to the beginning of the file, replacing `<YEAR>` and `<NAME>` with the current year and your name:

@@

Note that `Manuscript.qmd` was provided to you under [CC\ BY\ 4.0](https://creativecommons.org/licenses/by/4.0/), which is what you indicate with the previous comment. If you edited the file, you may also add yourself.

>>>>

::: {#cau-exclusive-rights .callout-caution}
### Giving Publishers Exclusive Rights

When you have your article published under an open access license, some publishers still demand an exclusive publishing and distribution license that goes further than the open access license available to readers of the journal [@Rumsey2022]. However, technically this is not required for publishers to provide their services [@Suber2022] and authors should oppose this practice. If your chosen publisher insists on an exclusive license, you may at least retain the copyright for your figures -- follow the guide "Retaining copyright for figures in academic publications to allow easy citation and reuse" by @Elson2016 to learn how to do that.

@@

For example, the `apaquarto` extension that you included in your project is a work by others.^[Unless you are the author of the `apaquarto` extension.] You need to indicate its license so that others know what they are allowed to do -- and, of course, you need to comply with any terms yourself, such as retaining the copyright notice.^[In this particular case, `apaquarto` is licensed under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/), which does not require you to retain a copyright notice. In fact, it's an extremely permissive license.] In contrast, if it were not for this tutorial, the manuscript would contain your own work and you would need to indicate under which license you provide it to others.

@@

:::: {}
- [CC licenses](https://creativecommons.org/faq/#how-do-i-apply-a-creative-commons-license-to-my-material)
- [ODbL\ 1.0](https://opendatacommons.org/licenses/odbl/)
::::
:::

@@

In the simplest case, one just creates a file called `LICENSE.txt` where the full text of the license is copied verbatim. This is a practice propagated by GitHub, which provides instructions for many licenses via [ChooseALicense.com](https://choosealicense.com/licenses/). However, if the project is not completely covered by one single license, this practice may become unwieldy. For example, if a project contains different types of works by different authors, the `LICENSE.txt` needs to detail which file is covered by which license(s), along with any copyright notices.

@@

Every major free/open license has a unique [SPDX identifier](https://spdx.org/licenses/) which allows communicating the license choice unequivocally. We will be using that to indicate the license for every file in your project, along with the year of publication and the copyright holder. To do this, we add a comment to the beginning of every file and include the two tags `SPDX-FileCopyrightText` and `SPDX-License-Identifier`. How this works depends on the file type, as the syntax for a comment varies.

For example, if you previously created the file `create_data_dictionary.R`, you can now add the following comment to the beginning of the file, replacing `<YEAR>` and `<NAME>` with the current year and your name -- of course, you can also choose a different license:

@@

Note that `Manuscript.qmd` was provided to you under [CC\ BY\ 4.0](https://creativecommons.org/licenses/by/4.0/), which is what you indicate with the previous comment. If you edited the file, you may also add yourself.^[For licenses that require that modifications are indicated, this is an easy way to comply with them, although you do not need to provide your real name.]
